---
title: "LiveSessionUnit11"
author: "Jacob Brionez"
date: "November 20, 2016"
output: html_document
---

```{r}
library(fpp)
data("hsales")
```

## a)	Plot the time series. Can you identify seasonal fluctuations and/or a trend? 
Yes it can be seen that there are highs and lows at regular intervals indicating a seasonal cycle.

```{r}
plot(hsales)
```

## b)	Use a classical decomposition to calculate the trend-cycle and seasonal indices. 

```{r}
fitd <- decompose(hsales)
plot(fitd)
```

## c)	Do the results support the graphical interpretation from part (a)? 
Yes, it does show in the seasonal decomposition that there are similar patterns in the time scale.

```{r}

```

##d)	Compute and plot the seasonally adjusted data. 

```{r}

eeadj <- seasadj(fitd)
plot(eeadj)
```

## e)	Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier? 
The outlier causes a spike in the seasonally adjusted data and affected a period of time before and after the spike that looks to be about a year before and after the spike.

```{r}
hsales[55] <- hsales[55]+500
fitd2 <- decompose(hsales)
plot(fitd2)
eeadj2 <- seasadj(fitd2)
plot(eeadj2)
```

## f)	Does it make any difference if the outlier is near the end rather than in the middle of the time series? 
With the exception of the placement of the outlier it does not seem to make a difference in the overall time series upon visual inspection.

```{r}
hsales[55] <- hsales[55]-500
hsales[200] <- hsales[200]+500
fitd2 <- decompose(hsales)
plot(fitd2)
eeadj2 <- seasadj(fitd2)
plot(eeadj2)
```

##g)	Use STL to decompose the series. 

```{r}
fit <- stl(hsales, s.window=5)
plot(fit)
plot(hsales, col="gray",
main="Effect of Outlier",
ylab="hsales", xlab="")
lines(fit$time.series[,2],col="red",ylab="Trend")
```